---
output: 
  html_document: 
    highlight: monochrome
    number_sections: yes
    theme: flatly
    toc: yes
---
**Course Diary**

#Tools and methods for open and reproducible research

**Work done during the week and learning summary.**

 - Completed the course DataCamp assay: "R short and sweet"";

 - Completed also the optional "Introduction to R" course in DataCamp available to all users; 

 - Installed RStudio, GitHub desktop, and explored the RMarkdown Basics & R Code Chunks provided help lists.  

    - Succesfully forked the IODS project to personal repository & made own repository online link  

##Summary of what was accomplished this week.

 - Completed the following courses on DataCamp: _"R short and sweet"_  ; _"Introduction to R"_ & 1^st^ levels of _DataCamp (EXTRA): Helsinki Social Statistics_ 

 - Installed RStudio, GitHub desktop, and explored the RMarkdown Basics & R Code Chunks provided help lists.  

    - Succesfully forked the IODS project to personal repository & made own repository online link  
    
    - Created the new file RMarkdown (README.Rmd)

_**Learned how to extract information from data**_

Modifications that seemed adequate to the provided data base

```{r cars, echo = FALSE}
str(cars)
summary(cars)


```

_**Learned how to modify Plots**_

Shown the modifications that seemed adequate to the provided example: 

- Modify the data points colors 

- Insert plot title 

- Change x , y axis legends 


```{r pressure, echo=FALSE}
plot(pressure, xlab = "Temperature" , ylab = "Pressure" ,  col = "tomato", bg = "grey", main = "Temperature - Pressure curve")
```


