---
output: 
  html_document: 
    highlight: monochrome
    toc: yes
---
**Course Diary**

# Regression and model validation

*During this chapter the amount of information to be memorized is higher and more complex. The datacamp exercise, including the videos included in it touch the surface of what you can do to build graphs to visualize and analyse your data. Note that during this chapter you will definetily need to consult external help cheats (mostly provided during the datacamp and your mooc links)*


## **Work done during the week and learning summary.**

1. _**Data wrangling exercise**_ 

Load the data set provided (questionary answers regarding learning approaches and students accomplishments in a statistics course):
 
 - All data provided in the url 
 
```{r}
#read the data from URL provided link
lrn14 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS3-data.txt", sep="\t", header=TRUE)
```
  
 
 
Data after data wrangling exercises (stored locally)

 
```{r}
#read the data from file located in your hard drive
learning2014 <- read.csv("learning2014.csv")
```
 
   - Questions related to the following topics were placed together:     
        - deep_questions , surface_questions , strategic_questions 
        

2. _**How to extract information from data**_

Since the full data set can be overwelming to explore, viewing the structure, summary and the first values of the data can be helpful to identify what are your main interests in this questionary answers. 

```{r, echo = FALSE}
#view the structure and summary of the non filtered "learning2014" data
str(lrn14)
summary(lrn14)
head(lrn14)
```

- View the structure and summary of the filtered "learning2014" data

```{r, echo = FALSE}
#view the structure and summary of the filtered "learning2014" data
str(learning2014)
summary(learning2014)
head(learning2014)
```

A careful observation of the loaded data shows you that from all the 63 original variables I filtered 7 variables of interest (Gender , Age, Attitude, Points, Deep questions, Surface questions, Strategic questions). 


_**View data in Plots format**_

Graphs to visualize the overall relationships betwwen all the selected variables 

```{r, echo = FALSE}
#Draw all possible scatterplots to explore the data frame
library(GGally)
library(ggplot2)
p <- ggpairs(data = learning2014, mapping = aes(col = gender, alpha = 0.3), lower = list(combo = wrap("facethist", bins = 20)))
p
```

Focusing on the diagonal set of graphs (7 variables by gender):
You can already check that mainly in the Attitude variable the differences between gender are more noticeable; 

From all the variables analysed, strategic_questions is the one that shows a more normal distributon, all the other data seems slighly to dramatically skewed. 


- Based in the previous observations it seems interesting to visualize how the attitute (that seems to vary most with gender) correlates with the acquired Points. 

```{r pressure, echo=FALSE}
# plot with the 
p1 <- ggplot(data = learning2014, aes(x = Attitude, y = Points, col = gender))

# define the visualization type (points)
p2 <- p1 + geom_point()

# add a regression line
p3 <- p2 + geom_smooth(method = "lm")
p3

```

You can verify in the graph how the correlation trend varies between female (red line) and male (blue line). The grey area in the graph shows you how this data varies. 


```{r}
lm(formula = Points ~ Attitude + stra + surf, data = learning2014)
```

Analyses of the linear regression tell you that by each attitude point up the sutdents have a chance of getting mote 0.34 point , as for one level up of strategic questions 0.85 and inversevly correlated for the surface points losing -0.59

_**Graphical model validation**_



_**Residuals vs Fitted plot**_



_**Normal Q-Q plot**_

```{r}
my_fit <- lm(formula = Points ~ Attitude + stra + surf, data = learning2014)
```



