---
output: 
  html_document: 
    highlight: monochrome
    toc: yes
---
**Course Diary**

# CHAPTER 2 - Regression and model validation

*During this chapter the amount of information to be memorized is higher and more complex. The datacamp exercise, including the videos included in it touch the surface of what you can do to build graphs to visualize and analyse your data. Note that during this chapter you will definetily need to consult external help cheats (mostly provided during the datacamp and your mooc links)*


**Work done during the week and learning summary.**

Data wrangling exercise: 

1. Step is to load the data set you will work on:
 
 - Either from the provided url 
```{r, echo = FALSE}
# Read the data from URL provided link
lrn14 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS3-data.txt", sep="\t", header=TRUE)
```
  
 - Or from the file locally stored in your harddrive 
 
```{r}
# Read the data from file located in yout hard drive
read.csv("learning2014", header=TRUE, sep = "\t" , row.names = 1)
```
 



_**Learned how to extract information from data**_

Did the following modifications that seemed adequate to the provided data base


```{r, echo = FALSE}
#analyse "Cars" data
read.csv("learning2014.csv", header=TRUE, sep = "\t")
str(learning2014)
summary(learning2014)


```

_**Learned how to modify Plots**_

Shown the modifications that seemed adequate to the provided example: 

- Modify the data points colors 

- Insert plot title 

- Change x , y axis legends 

#Plot a curve to analyse the correlations between temprature and pressure
```{r pressure, echo=FALSE}
#Plot a curve to analyse the correlations between temprature and pressure
plot(pressure, xlab = "Temperature" , ylab = "Pressure" ,  col = "tomato", bg = "grey", main = "Temperature - Pressure curve")
```


