---
output: 
  html_document: 
    highlight: monochrome
    toc: yes
---
**Course Diary**

# CHAPTER 2 - Regression and model validation

*During this chapter the amount of information to be memorized is higher and more complex. The datacamp exercise, including the videos included in it touch the surface of what you can do to build graphs to visualize and analyse your data. Note that during this chapter you will definetily need to consult external help cheats (mostly provided during the datacamp and your mooc links)*


**Work done during the week and learning summary.**

1. _**Data wrangling exercise:**_ 

Load the data set:
 
 - All data provided in the url 
 
```{r, echo = FALSE}
#read the data from URL provided link
lrn14 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS3-data.txt", sep="\t", header=TRUE)
```
  
 - Data after data wrangling exercises (stored locally)
 
```{r, echo = FALSE}
#read the data from file located in your hard drive
learning2014 <- read.csv("learning2014.csv")
```
 
   - Questions related to the following topics were placed together:     
        - deep_questions , surface_questions , strategic_questions 
        

2. _**How to extract information from data:**_

Since the full data set can be overwelming to explore, viewing the structure, summary and the first values of the data can be helpful to identify what are your main interests in this questionary answers. 

```{r, echo = FALSE}
#view the structure and summary of the non filtered "learning2014" data
str(lrn14)
summary(lrn14)
head(lrn14)
```


```{r, echo = FALSE}
#view the structure and summary of the non filtered "learning2014" data
str(learning2014)
summary(learning2014)
head(learning2014)
```

A careful observation of the loaded data shows you that from all the 63 original variables I filtered 7 variables of interest (Gender , Age, Attitude, Points, Deep questions, Surface questions, Strategic questions). 


_**View data in Plots format**_

Shown the modifications that seemed adequate to the provided example: 

- Modify the data points colors 

- Insert plot title 

- Change x , y axis legends 

#Plot a curve to analyse the correlations between temprature and pressure
```{r pressure, echo=FALSE}
#Plot a curve to analyse the correlations between temprature and pressure
plot(data = lrn14, attitude, points, aes(xlab = "Attitude" , ylab = "Points" ,  col = "tomato", bg = "grey", main = "Points Vs Attitude""))
```

library(ggplot2)

qplot(attitude, points, data = learning2014) + geom_smooth(method = "lm")
