---
title: "Course Diary"
author: "**Elsa Marques**"
date: "1 November 2017"
output:
  html_document:
    highlight: monochrome
    number_sections: yes
    theme: flatly
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**First impressions about the course**

*At first **overwhelming** for non experienced programmers or advanced users in the computer field. Nonetheless, after a few hours of careful reading instructions and self learning sessions in the provided tools one can start to understand that **sky is the limit** for the stubborn and motivated individuals.*

_**Elsa Marques repository link:**_

<https://github.com/ElsaMarques/IODS-project.git>

_**Course diary link:_** 

<https://ElsaMarques.github.io/IODS-project/>


#Tools and methods for open and reproducible research

##Summary of what was accomplished this week.

 - Completed the following courses on DataCamp: _"R short and sweet"_  ; _"Introduction to R"_ & 1^st^ levels of _DataCamp (EXTRA): Helsinki Social Statistics_ 

 - Installed RStudio, GitHub desktop, and explored the RMarkdown Basics & R Code Chunks provided help lists.  

    - Succesfully forked the IODS project to personal repository & made own repository online link  
    
    - Created the new file RMarkdown (README.Rmd)

_**Learned how to extract information from data**_

Modifications that seemed adequate to the provided data base

```{r cars, echo = FALSE}
str(cars)
summary(cars)


```

_**Learned how to modify Plots**_

Shown the modifications that seemed adequate to the provided example: 

- Modify the data points colors 

- Insert plot title 

- Change x , y axis legends 


```{r pressure, echo=FALSE}
plot(pressure, xlab = "Temperature" , ylab = "Pressure" ,  col = "tomato", bg = "grey", main = "Temperature - Pressure curve")
```

#Regression and model validation
#Logistic regression
#Clustering and classification
#Dimensionality reduction techniques
#Final assignment